---
title: "problem set 1"
author: "Chandler James"
date: "October 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# library loadings
library(tidyverse)
library(haven)
library(stargazer)
```

# Part 1

```{r}
# load the data 

pres_polls_2016 <- read_dta("~/Box Sync/SOC/data/pres_polls_2016.dta")


```

### 1

```{r}
# generate a variable, dem_pollshare, measuring the two party share of support for the Democrats based on the variables adjpoll_clinton and adjpoll_trump

pres_polls_2016 <- pres_polls_2016 %>%
  mutate(dem_pollshare = adjpoll_clinton/(adjpoll_clinton + adjpoll_trump))
```

### 2
```{r}
# generate a variable, demvoteshare, measuring the Democratic Party’s share of thetwo-party vote based on the variables DemVotes and RepVotes.

pres_polls_2016 <- pres_polls_2016 %>%
  mutate(demvoteshare = DemVotes/(DemVotes + RepVotes))
```

### 3
```{r}
# generate a variable, pollsminusvotes, measuring the difference between the two vari-ables (dempollshare minus demvoteshare)

pres_polls_2016 <- pres_polls_2016 %>%
  mutate(pollsminusvotes = dem_pollshare-demvoteshare)
```

### 4
```{r}
# generate a variable, abspollsminusvotes, measuring the absolute value of the differ-ence between the two variables (dempollshare minus demvoteshare).

pres_polls_2016 <- pres_polls_2016 %>%
  mutate(abspollsminusvotes = abs(dem_pollshare - demvoteshare))
```


# Part 2

### 1
```{r}
# Make a scatter plot with dempollshare on the x-axis and demvoteshare on the y-axis.

pres_polls_2016 %>%
  ggplot(aes(dem_pollshare, demvoteshare)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### 2
The graph above shows considerable correspondence between the x-axis (dem_pollshare) and y-axis (dem_voteshare). That is to say that the independent variable predicts the dependent variable. The slope of the line is near 45 degrees indicating that x predicts y reasonably well.

# Part 3

```{r}
summary(pres_polls_2016)
```

### 1
around 30.3% of the polls are considered to be top quality.


```{r}
# filtering for the top quality polls
top_pres_polls_2016 <- pres_polls_2016 %>%
  filter(quality_a == 1)
summary(top_pres_polls_2016)

# filtering for the low quality polls

low_pres_polls_2016 <- pres_polls_2016 %>%
  filter(quality_a == 0)
summary(low_pres_polls_2016)
```
### 2
The mean of abs_poll_minus_votes for top quality polls is  0.0333.
The mean of abs_poll_minus_votes for low quality polls is  0.0399.

### 3 

The results above suggest that the top quality polls are more accurate. That is because the poll is closer to the vote when the value of abs_poll_minus_votes is smaller.

# Part 4: Sample Size

```{r}
mod1 <- lm(abspollsminusvotes ~ samplesize, data = pres_polls_2016)
stargazer(mod1, type = "text")
```

The coefficient on the variable sample size is negative. This suggest that there is a negative relationship between sample size and distance between the poll and the vote. This is evidence to suggest that larger sample sizes does not necessarily improve poll accuracy. 










